\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-11-11 archive.tex
working on archive

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{dblfloatfix}    % To enable figures at the bottom of page
% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{booktabs} \usepackage{stmaryrd}% for professional tables
\usepackage{stmaryrd}(
\usepackage[section]{placeins}
% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
%\usepackage{hyperref}

@@

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{tikz}
\usepackage{circuitikz}
\usetikzlibrary{fit,bayesnet}
%\usetikzlibrary{arrows.meta}
%\usetikzlibrary{positioning}
%\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\tikzset{squiggle/.style={decorate, decoration={snake,amplitude=.4mm}}}
\usepackage{dsfont}
\usepackage{amsmath}

@@

The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

Using the frontiers we  define the following intuitive lower bound on the joint probability, called $\lowerBound$:
\begin{align}
 J\geq \lowerBound\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
\end{align}
We alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$  (\textbf{Wake})
and $(\mathcal{D},\theta)$ (\textbf{Sleep-G}):

\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]\propto\probability[p|x,\mathcal{D},\theta]$ 
is large (i.e., programs with high posterior probability).
%Here we use a simple and generic enumerative program synthesis algorithm.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{Sleep-G: Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
and then updating $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.
%To represent DSLs and search the space of DSLs, we use Fragment Grammars~\cite{tim}.

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
during the \textbf{Sleep-R} phase: $q$ is trained to approximate the
posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}}]$.
  Thus training the neural network amortizes the cost of finding programs with high posterior probability.

\textbf{Sleep-R: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$q(p|x)$ to assign high probability to programs $p$ where
$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because incorporating those programs
into the frontiers will most increase $\lowerBound$.

Intuitively, this 3-phase inference procedure can work in practice because each of the 3 phases bootstraps off of the others (Figure~\ref{feeding}). As the DSL grows and as the recognition model becomes more accurate, waking becomes more effective, allowing the agent to solve more tasks; when we solve more tasks during waking,
the Sleep-G phase has more data from which to learn the DSL; and, because Sleep-R is trained on both
samples from the DSL and programs found during waking,
the recognition model gets both more data, and higher-quality data, whenever the DSL improves and whenever we discover more successful programs.
\begin{figure}\centering
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $q$};

@@

>>>>

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{dblfloatfix}    % To enable figures at the bottom of page
% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{booktabs} \usepackage{stmaryrd}% for professional tables

\usepackage[section]{placeins}
% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
%\usepackage{hyperref}

@@

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{tikz}
\usepackage{circuitikz}
\usetikzlibrary{fit,bayesnet,calc}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{positioning}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\tikzset{squiggle/.style={decorate, decoration={snake,amplitude=.4mm}}}
\usepackage{dsfont}
\usepackage{amsmath}

@@

The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
\begin{definition}
   A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
  \end{definition}
Using the frontiers we  define the following intuitive lower bound on the joint probability, called $\lowerBound$:
\begin{align}
 J\geq \lowerBound\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
\end{align}
We alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$  (\textbf{Wake})
and $(\mathcal{D},\theta)$ (\textbf{Sleep-G}):
\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]\propto\probability[p|x,\mathcal{D},\theta]$ 
is large (i.e., programs with high posterior probability).
%Here we use a simple and generic enumerative program synthesis algorithm.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{Sleep-G: Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
and then updating $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.
%To represent DSLs and search the space of DSLs, we use Fragment Grammars~\cite{tim}.

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $Q(\cdot |\cdot )$,
during the \textbf{Sleep-R} phase: $Q$ is trained to approximate the
posterior over programs, $Q(p|x)\approx \probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}]$.
  Thus training the neural network amortizes the cost of finding programs with high posterior probability.

\noindent\textbf{Sleep-R: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$Q(p|x)$ to assign high probability to programs $p$ where
$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because incorporating those programs
into the frontiers will most increase $\lowerBound$.

Intuitively, this 3-phase inference procedure can work in practice because each of the 3 phases bootstraps off of the others (Figure~\ref{feeding}). As the DSL grows and as the recognition model becomes more accurate, waking becomes more effective, allowing the agent to solve more tasks; when we solve more tasks during waking,
the Sleep-G phase has more data from which to learn the DSL; and, because Sleep-R is trained on both
samples from the DSL and programs found during waking,
the recognition model gets both more data, and higher-quality data, whenever the DSL improves and whenever we discover more successful programs.
\begin{figure}\centering
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};

@@

\begin{definition}
A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
one number for each DSL element $e\in \mathcal{D}$, written $\theta_e$ and controlling the probability of  $e$ occurring in a program,
and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.
\end{definition}

Together with its weight vector, a DSL defines a distribution over
programs, $\probability[p|\mathcal{D},\theta]$.  We define this
distribution by specifying a procedure for drawing samples from
$\probability[p|\mathcal{D},\theta]$ (Algorithm~\ref{sampleProgram}).
Care must be taken to ensure that variable scoping rules are obeyed
and that programs are well-typed.  We ensure well-typed programs by
performing Hindley-Milner type inference~\cite{pierce} during
sampling, and assume that each task is annotated with the type of the
program that will solve it.
 \begin{algorithm}
   \caption{Generative model over programs}
   \label{sampleProgram}
   \begin{algorithmic}[1]
     \State \textbf{function} sample$(\mathcal{D}, \theta, \tau)$:
     \State {\bfseries Input:} DSL $(\mathcal{D},\theta)$, type $\tau$
     \State \textbf{Output:} a program whose type unifies with $\tau$
     \State  \textbf{return }sample$'(\mathcal{D}, \theta, \varnothing, \tau)$

     \Statex
     \State \textbf{function} sample$'(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
     \State {\bfseries Input:} DSL $(\mathcal{D},\theta)$, environment $\mathcal{E}$, type $\tau$\Comment{Environment $\mathcal{E}$ starts out as $\varnothing$}
     \State \textbf{Output:} a program whose type unifies with $\tau$
     \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
     \State var $\gets$ an unused variable name
     \State body $\sim$ sample$'(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$\Comment{Recursively sample function body}
     \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
     %\Endif
     \Else\Comment{Build an application to give something w/ type $\tau$}
     \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
     $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
     \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
     \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
       \theta_e&\text{ if }e\in \mathcal{D}\\
\theta_{var}/|\text{variables}|&\text{ if }e\in \mathcal{E}
       \end{cases}$
     \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
     \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
     %   \State unify$(\tau,\beta)$
     \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
     \State $a_k\sim\text{sample}'(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
     \EndFor
     \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
     \EndIf
     \Statex
     \Statex\textbf{where:}
     \State yield$(\tau) = \begin{cases}
       \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
       \tau   &\text{ otherwise.}
     \end{cases}$ \Comment{Final return type of $\tau$}
     \State  args$(\tau) = \begin{cases}
       [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
       []   &\text{ otherwise.}
     \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
   \end{algorithmic}
 \end{algorithm}

The purpose of training the recognition model is to amortize the cost
of searching for programs.  It does this by learning to predict, for
each task, programs with high likelihood according to
$\probability[x|p]$ while also being probable under the prior
$(\mathcal{D},\theta)$,
thereby approximating the posterior $\probability[\cdot |x,\mathcal{D},\theta]$.

\subsubsection{Training $Q$}\label{recognitionTraining}
How should we get the data to train $Q$?
This is non-obvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from task/program pairs).
One approach is to draw programs from the DSL,
run them to get their input/outputs,
and then train $Q$ to predict the program from the input/outputs.
This is like how the wake-sleep algorithm for the Helmholtz machine
trains its recognition model during its sleep phase~\cite{dayan1995helmholtz}.
The advantage of training on samples, or ``dreams,'' is that
we can draw unlimited samples from the DSL,
training on a large amount of data.
Another approach is
to train $Q$ on the (program, task)
pairs discovered by the Wake phase.
The advantage here
is that the training data is much higher quality,
because we are training on real tasks.
Due to these complementary advantages,
we train on both these sources of data.

Formally, $Q$ should approximate the true posteriors over programs.
We can either train $Q$ to perform full posterior inference by minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|Q(p|x) \right) \right]$,
or we can train $Q$ to perform MAP inference
by maximizing $\expect\left[\max_{p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]} \log Q(p|x) \right]$,
where in both cases the expectation is taken over tasks. Taking this expectation over the empirical distribution of tasks trains $Q$ on the real data; taking it over samples from the generative model trains $Q$ on ``dreams.''
We define a pair of alternative objectives for the recognition model,
$\mathcal{L}_{\text{RM}}^{\text{posterior}}$ and $\mathcal{L}_{\text{RM}}^{\text{MAP}}$,
which either train $Q$ to perform full posterior inference or MAP inference, respectively.
These objectives combine real-data and dream data:
\begin{align*}
  \mathcal{L}_{\text{RM}}^{\text{posterior}} &= \mathcal{L}_{\text{Real}}^{\text{posterior}} + \mathcal{L}_{\text{Dream}}^{\text{posterior}}&
  \mathcal{L}_{\text{RM}}^{\text{MAP}} &= \mathcal{L}_{\text{Real}}^{\text{MAP}} + \mathcal{L}_{\text{Dream}}^{\text{MAP}}\\
  \mathcal{L}_{\text{Real}}^{\text{posterior}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
    \frac{\probability\left[x,p|\mathcal{D},\theta \right]\log Q(p|x)}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\right] &
\mathcal{L}_{\text{Real}}^{\text{MAP}}& = \expect_{x\sim X}\left[\max_{\substack{p\in \mathcal{F}_x\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}} \log Q(p|x) \right]  \\
  \mathcal{L}_{\text{Dream}}^{\text{posterior}} &= \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log Q(p|x)\right]&
    \mathcal{L}_{\text{Dream}}^{\text{MAP}} &= \expect_{x\sim(\mathcal{D},\theta) }\left[\max_{\substack{p\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}}\log Q(p)\right]
\end{align*}
  The `dream' objectives are essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.
But, evaluating $\mathcal{L}_{\text{Dream}}$ involves drawing programs from
the current DSL, running them to get their outputs,
and then training $Q$ to regress from the input/outputs to the program.
Since these programs map inputs to outputs,
we need to sample the inputs as well.
Our solution is to sample the inputs
from the empirical observed distribution of inputs in $X$.

The $\mathcal{L}_{\text{Dream}}^{\text{MAP}}$ objective involves
finding the MAP program solving a task drawn from the DSL.
To make this tractable,
rather than \emph{sample} programs as training data for $\mathcal{L}_{\text{Dream}}^{\text{MAP}}$,
we \emph{enumerate} programs
in decreasing order of their prior probability,
tracking,
for each dreamed task $x$,
the set of enumerated programs maximizing
$\probability[x,p|\mathcal{D},\theta]$.

\subsubsection{Parameterizing $Q$}\label{recognitionParameterization}

Broadly the literature contains two different approaches to
parameterizing conditional distributions over programs.  The first
approach, seen in systems like RobustFill~\cite{devlin2017robustfill}
and~\cite{zavershynskyi2018naps}, is to use a recurrent neural network
to predict the program token-by-token as in Seq2Seq or Seq2Tree.  The
advantage of this approach is that the neural net can predict the
entire program, so if the network is sufficiently powerful, it can
completely solve the synthesis problem.  There are two disadvantages
to this approach.  First, these models can perform poorly at out-of-sample
generalization~\cite{}, which is critical for our setting, as the
agent may need to solve new tasks that are qualitatively different
from the tasks it has solved so far.  Second, a powerful deep
recurrent network may be costly to sample or enumerate from ---
so if the network cannot easily solve a task, we cannot compensate
with rapid sampling or enumeration.  In contrast, state-of-the-art
enumerative program synthesizers evaluate millions of programs per
second~\cite{feser2015synthesizing}.

We adopt a middle ground between these two extremes.  Our recognition
model predicts a distribution over primitives in the DSL,
conditioned on the local context in the syntax tree of the
program. When predicting the next node to add to the syntax tree of a program,
the recognition model
conditions on the parent node, as well as
which argument is being generated.
This is a kind of `bigram' model over trees,
where the bigrams
take the form of (parent, child, argument index).

Formally,
$Q$ predicts a $(|\mathcal{D}|+2)\times (|\mathcal{D}|+1)\times A$-dimensional tensor,
where $A$ is the maximum arity\footnote{The arity of a function is the number of arguments that it takes as input.} of any primitive in the DSL.
Slightly abusing notation, we write this tensor as $Q_{ijk}(x)$,
where $x$ is a task,
$i\in \mathcal{D}\cup\left\{\text{start},\text{var}\right\}$,
$j\in \mathcal{D}\cup\left\{\text{var} \right\}$,
and $k\in \left\{1,2,\cdots,A \right\}$.
The output $Q_{ijk}(x)$
controls the probability of
sampling primitive $j$ given that
$i$ is the parent node in the syntax tree
and we are sampling the $k^{\text{th}}$ argument.
Figure~\ref{bg} diagrams this generative process, Algorithm~\ref{recognitionSample}
specifies a sampling procedure for $Q(\cdot |x)$,
and Figure~\ref{neuralPipeline} diagrams the neurally guided program inference procedure.
This parameterization
has three main advantages:
(1) it supports fast enumeration
and sampling of programs,
because the
recognition model
only needs to run once for each task;
(2) it allows the recognition model to provide fine-grained
information about the structure of the target program;
and (3)
training this recognition model
causes it to learn to break symmetries in the space of programs,
described next.
\begin{figure}
\centering  \begin{tikzpicture}%[every node/.style={inner sep=0,outer sep=0}]
    \node(l)[draw] at (0,0) {\texttt{$\lambda$ (a)}};
    \node(k)[draw] at ([yshift=-50]l.south) {\texttt{+}};
    \node(o)[draw] at ([xshift=-50,yshift=-50]k.south) {\texttt{9}};
    \node(m)[draw] at ([xshift=50,yshift=-50]k.south) {\texttt{*}};
    \node(x1)[draw] at ([xshift=50,yshift=-50]m.south) {\texttt{a}};
    \node(x2)[draw] at ([xshift=-50,yshift=-50]m.south) {\texttt{a}};

    \draw[->] (l.south)-- node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\text{start},\texttt{+},1}(x)$}(k.north);
    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{9},1}(x)$}(o.north);
    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{*},2}(x)$}(m.north);
    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},2}(x)$}(x1.north);
    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},1}(x)$}(x2.north);
\end{tikzpicture}
\caption{Parameterization of distribution over programs predicted by recognition model.
  Here the program (syntax tree shown above) is \texttt{($\lambda$ (a) (+ 9 (* a a )))}.
Each conditional distribution predicted by the recognition model is written $Q_{\text{parent},\text{child},\text{argument index}}(x)$, where $x$ is a task.}\label{bg}
\end{figure}
 \begin{algorithm}
   \caption{Drawing from distribution over programs predicted by recognition model. Compare w/ Algorithm~\ref{sampleProgram}}
   \label{recognitionSample}
   \begin{algorithmic}[1]
     \State \textbf{function} recognitionSample$(Q, x, \mathcal{D}, \tau)$:
     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, type $\tau$
     \State \textbf{Output:} a program whose type unifies with $\tau$
     \State \textbf{return } $\text{recognitionSample}'(Q,x,\text{start},1,\mathcal{D},\varnothing,\tau)$
     \Statex

     \State \textbf{function} recognitionSample$'(Q, x, \text{parent}, \text{argumentIndex}, \mathcal{D}, \mathcal{E}, \tau)$:
     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, parent $\in \mathcal{D}\cup\left\{\text{start},\text{var} \right\}$, argumentIndex $\in \mathbb{N}$, environment $\mathcal{E}$, type $\tau$
     \State \textbf{Output:} a program whose type unifies with $\tau$
     \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
     \State var $\gets$ an unused variable name
     \State body $\sim$ recognitionSample$'(Q,x,\text{parent},\text{argumentIndex},\mathcal{D},\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$%\Comment{Recursively sample function body}
     \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
     %\Endif
     \Else\Comment{Build an application to give something w/ type $\tau$}
     \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
     $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
     \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
     \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
       Q_{\text{parent},e,\text{argumentIndex}}(x)&\text{ if }e\in \mathcal{D}\\
Q_{\text{parent},\text{var},\text{argumentIndex}}(x)/|\text{variables}|&\text{ if }e\in \mathcal{E}
       \end{cases}$
     \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
     \State newParent$\gets\begin{cases}
     e &\text{ if }e\in \mathcal{D}\\
     \text{var}&\text{ if }e\in \mathcal{E}\end{cases}$
     \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
     %   \State unify$(\tau,\beta)$
     \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
     \State $a_k\sim\text{recognitionSample}'(Q,x,\text{newParent},k,\mathcal{D},\mathcal{E},\alpha_k)$
     \EndFor
     \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
     \EndIf
     %% \Statex
     %% \Statex\textbf{where:}
     %% \State yield$(\tau) = \begin{cases}
     %%   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
     %%   \tau   &\text{ otherwise.}
     %% \end{cases}$ \Comment{Final return type of $\tau$}
     %% \State  args$(\tau) = \begin{cases}
     %%   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
     %%   []   &\text{ otherwise.}
     %% \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
   \end{algorithmic}
 \end{algorithm}
 \begin{figure}
   \begin{tikzpicture}
     \node(x)[draw,align=center] at (0,0) {Task $x$\\e.g., input/outputs};
     \node(e)[draw, align=center] at ([xshift=3cm]x.east) {Domain-specific encoder\\
       e.g., RNN or CNN};
     \node(m)[draw,align = center] at ([xshift=1.1cm]e.east) {MLP};%\\Output: $(|\mathcal{D}| + 2)\times(|\mathcal{D}| + 1)\times A$};
     \node[draw,align = center](s) at ([xshift=3cm]m.east) {Enumerate\\(Algorithm~\ref{recognitionSample})};
     \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Frontier\\$\mathcal{F}_x$};

     \draw[->] (x.east) -- (e.west);
     \draw[->] (e.east) -- (m.west);
     \draw[->] (m.east) -- node(p)[fill=white,align=center,midway,inner sep=0,outer sep=0,rotate=90]{\small $Q_{ijk}(x)$}(s.west);
     \draw[->] (s.east) -- (f.west);
     \draw[->] (x.north) -- ([yshift=15]x.north) -- node[above = 0pt,midway]{\small likelihood $\probability[x|p]$}([yshift=15,xshift=40]s.north) -- ($(s.east)!0.5!(f.west)$);
     \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
     ([xshift = -5,yshift = -25]e.west) -- node[below = 3pt] {{Recognition model $Q(p|x)$}} ([xshift = 155,yshift = -25]e.west);
     \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
     ([xshift = -5,yshift = -25]s.west) -- node[below = 3pt] {{ Search}} ([xshift = 5,yshift = -25]f.east);
     %% \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
     %% ([xshift = -5,yshift = -25]p.west) -- node[below = 3pt] {{\small Params}} ([xshift = 5,yshift = -25]p.east);
   \end{tikzpicture}
   \caption{Neurally-guided program inference pipeline. Recognition model outputs distribution over program $Q(p|x)$ parameterized by tensor $Q_{ijk}$. Program output by enumerative search incorporated into frontier if likelihood $\probability[x|p] > 0$}
   \label{neuralPipeline}
   \end{figure}

\subsubsection{Learning to break symmetries in program space}

A good DSL not only exposes high-level building blocks,
but also carefully restricts the ways in which those building blocks are allowed to compose.
For example,
a DSL for list manipulation should contain both the empty list and
a routine for appending lists,
but should not allow
appending the empty list.
Similarly a DSL for arithmetic should contain both addition and
the number zero but
disallow adding zero.
These restrictions break symmetries in the space of programs.
Here we show that the
combination of the $\mathcal{L}^{\text{MAP}}$ training objective (Section~\ref{recognitionTraining}) and bigram parameterization (Section~\ref{recognitionParameterization})
interact in a way that spontaneously breaks symmetries in the program space,
allowing the agent to more efficiently
explore the space of programs.

Consider an agent tasked with writing programs built from addition, multiplication, and the constants zero and one.
A bigram parameterization of $Q$ allows
it to represent the fact that it should never add zero ($Q_{\code{+},\code{0},0} = Q_{\code{+},\code{0},1} = 0$)
or never multiply by  1 ($Q_{\code{*},\code{1},0} = Q_{\code{*},\code{1},1} = 0$), or
that addition should always associate to the right
($Q_{\code{+},\code{+},0} = 0$).
The $\mathcal{L}^{\text{MAP}}$ training objective encourages
learning these canonical forms.
Consider two recognition models, $Q_1$ and $Q_2$,
and two programs in frontier $\mathcal{F}_x$,
$p_1=\code{(+ (+ 1 1) 1)}$ and $p_2=\code{(+ 1 (+ 1 1))}$,
where
\begin{align*}
  Q_1(p_1|x) = \frac{\epsilon}{2}&\qquad Q_1(p_2|x) = \frac{\epsilon}{2}\\
  Q_2(p_1|x) = 0&\qquad  Q_2(p_2|x) = \epsilon
\end{align*}
i.e., $Q_2$ breaks a symmetry by forcing right associative addition,
but $Q_1$ does not, instead splitting its probability mass equally between $p_1$ and $p_2$.
Now because $\probability[p_1|\mathcal{D},\theta] = \probability[p_2|\mathcal{D},\theta]$
(Algorithm~\ref{sampleProgram}),
we have
\begin{align*}
  \mathcal{L}^{\text{posterior}}_{\text{real}}(Q_1)& = \frac{\probability[p_1|\mathcal{D},\theta]\log \frac{\epsilon}{2} + \probability[p_2|\mathcal{D},\theta]\log \frac{\epsilon}{2}}{\probability[p_1|\mathcal{D},\theta] + \probability[p_2|\mathcal{D},\theta]} = \log \frac{\epsilon}{2}\\
  \mathcal{L}^{\text{posterior}}_{\text{real}}(Q_2)& = \frac{\probability[p_1|\mathcal{D},\theta]\log 0 + \probability[p_2|\mathcal{D},\theta]\log \epsilon}{\probability[p_1|\mathcal{D},\theta] + \probability[p_2|\mathcal{D},\theta]} = +\infty\\
  \mathcal{L}^{\text{MAP}}_{\text{real}}(Q_1)& = \log Q_1(p_1)           = \log Q_1(p_2)   = \log \frac{\epsilon}{2}\\
  \mathcal{L}^{\text{MAP}}_{\text{real}}(Q_2)& = \log Q_2(p_2) \phantom{ = \log Q_1(p_2) } = \log \epsilon\\
\end{align*}
So the $\mathcal{L}^{\text{MAP}}$ prefers $Q_2$ (the symmetry breaking recognition model),
while $\mathcal{L}^{\text{posterior}}$ reverses this preference.
We can confirm this experimentally by training
four recognition models:
a

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
Intuitively, we want the algorithm to look at the programs found
during waking and generalize beyond them, both so the DSL can better
express the current solutions, and also so that the DSL might expose
new abstractions which will later be used to discover more programs.
Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{equation}
        \log \probability[\mathcal{D}] + 
\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
\label{AIC}
  \end{equation}
At a high level, our approach is to search locally through the space 
of DSLs, proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC}
fails to increase.  These small changes consist of introducing new
candidate $\lambda$-expressions into the DSL.

Formally, we want a DSL
which puts high probability on
any refactoring of the programs found during waking, maximizing:
\begin{equation}
        \log \probability[\mathcal{D}] + 
\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\manyReduce p}\probability[p'|\mathcal{D},\theta]\Bigg)
\label{factorObjective}
  \end{equation}
where $p'\manyReduce p$
is the standard notation for ``expression $p'$ evaluates to $p$ by the rules of $\lambda$-calculus''~\cite{pierce}.
